---
title: "Linear Regression Coding Assignment-1"
editor_options:
  chunk_output_type: inline
output: html_notebook
---

```{r}
# Load essential libraries
library(ggplot2)
library(dplyr)
library(HSAUR)
library(ggcorrplot)
```

```{r}
# Load the heptathlon dataset
data(heptathlon)
str(heptathlon)
```

```{r}
# Introduce a new column called sprint highlighting slow and fast sprinters
heptathlon = heptathlon %>% mutate(sprint = ifelse(run200m <= 25 & run800m <= 129, 'fast', 'slow'))
str(heptathlon)
```

```{r}
# Change sprint column to factor type
heptathlon['sprint'] = lapply(heptathlon['sprint'], as.factor)
str(heptathlon)

```

```{r}
# Make a scatter plot between *run200m* (x-axis) and *longjump* (y-axis). What do you observe from this plot?
p=ggplot(data=heptathlon, aes(x = run200m, y = longjump)) +
 geom_point(color = "blue") +
 labs(title = "Scatter Plot Example", x = "run200m", y = "Long jump")
p
```
## It is observed that few people were able to finish their race around 25-26 secs and jump around 6.0 m whereas it was also observed that one finished the race in less than 23 secs and was able to jump more than 7m.

```{r}
# Correlation between all pairs of continuous predictors (leave out sprint and the response variable score). What do you observe?
cor_matrix = cor(heptathlon %>% select(-c(sprint, score)))
ggcorrplot(cor_matrix, method = 'circle', lab = TRUE)

```


```{r}
# Make a scatter plot between *run200m* (x-axis) and *longjump* (y-axis) now with the data points color-coded using *sprint*. What do you observe from this plot?

p = ggplot(data = heptathlon) +
 geom_point(aes(x = run200m , y = longjump , color = sprint)) +
 labs(x = 'run200 (sec)', y = 'longjump (m)') +
 theme(axis.text = element_text(size = 12),
 axis.text.x = element_text(size = 14),
 axis.text.y = element_text(size = 14),
 axis.title = element_text(size = 14, face = "bold")) 
p

```
## we were able to observe that ones who completedthe race in 25 sec were slow runners and the fast runners who were able to complete in less time
```{r}
# Calculate Pearson's correlation between *run200m* and *longjump*. What do you observe?
cor(heptathlon$run200m, heptathlon$longjump)

```

```{r}
# How many levels does the categorical variable *sprint* have? What is the reference level? 
contrasts(heptathlon$sprint)
```
## There are 2 levels. Reference level is fast.

```{r}
# Fit a linear model for approximating *score* as a function of *sprint*. Print the model's summary. How accurate is the model? How do the slow athletes' scores compare to the fast ones?
model = lm(data = heptathlon, score~sprint)
summary(model)
```
```{r}
mean_slow = mean(heptathlon[heptathlon$sprint == 'slow', 'score'])
mean_fast = mean(heptathlon[heptathlon$sprint == 'fast','score'])
mean_slow-mean_fast

```
## R-squared value is around 40%.


```{r}
# Fit a linear model for approximating *score* as a function of *shot* and *sprint*. Print the model's summary and answer the following questions:

# 1. Did the addition of the new predictor *shot* improve the model accuracy? 
# 2. *True/false* (explain in one line): the model suggests that there is a possible linear relationship between an athlete's score and shotput performance.
# 3. For a 1 metre increase in shot put throw and with the same sprint performance, we can say with 95% confidence that the athlete's score will increase/decrease by an amount in the interval [?, ?].
model = lm(data = heptathlon, score~shot+sprint)
summary(model)

```
## 1. Yes the accuracy got increased to 67%.
## 2. True, because the p value is very small.
## 3. [249.7-1.9658,249.7+1.9658]


```{r}
#  Using the model built above, extract the slope and intercept for estimating the *score* of *slow* and *fast* athletes. 
intercept_slow = coef(model)["(Intercept)"] + coef(model)["sprintslow"]
intercept_fast = coef(model)["(Intercept)"] 
slope_slow = coef(model)["sprintslow"]
print(intercept_slow)

```
```{r}
print(intercept_fast)
```
```{r}
print(slope_slow)
```



```{r}
# Complete the code below to build a linear model for approximating *score* as a function of *shot* and *sprint* using the training data. Predict the model performance by applying it to the test data.
# Split the data into 80% train and 20% test parts

set.seed(0)
train_ind <- sample(nrow(heptathlon), size = 0.8 * nrow(heptathlon)) # 80% for training, assuming hData contains your data
hDataTrain <- heptathlon[train_ind, ]
hDataTest <- heptathlon[-train_ind, ]
# Build linear regression model
model <- lm(score ~ shot + sprint, data = hDataTrain)
# Predict on the test data
predictions <- predict(model, newdata = hDataTest)
# Print the true and predicted scores for the test data
print(data.frame(True_Score = hDataTest$score, Predicted_Score = predictions))



```
```{r}
# Calculate the model error (mean-squared error for test data)
mse <- mean((hDataTest$score - predictions)^2)
print(paste("Mean Squared Error:", mse))
```



```{r}
# Fit a linear model for approximating *score* as a function of *shot*, *javelin*, and *sprint*. Print the model's summary and answer the following questions:

#1. Did the addition of the new predictor *javelin* improve the model accuracy? 
#2. *True/false* (explain in one line): the model suggests that there is a possible linear relationship between an athlete's score and javelin performance.
#3. For a 1 metre increase in shot put throw and with the same javelin and sprint performance, we can say with 95% confidence that the athlete's score will increase/decrease by an amount in the interval [?, ?].
model = lm(data = heptathlon, score~shot+sprint+javelin)
summary(model)
```
## No.. Adding additional predictor didn’t increase the accuracy of the model
## No linear relationship between javelin and score because of the large p value.
## Interval [249-1.9660,249+1.9660]


```{r}
# Fit a linear model for approximating *score* as a function of *highjump*, and *sprint*. Print the model's summary and answer the following questions:
# 1. How accurate is this model?
# 2. Considering a p-value of 10% as cutoff, are there any insignificant features?

model=lm(data = heptathlon, score~highjump+sprint)
summary(model)

```
## Accuracy is around 83%
## By considering p values there are no insignificant features

